---
title: "Lab 07 - Conveying the right message through visualisation"
author: "George Nesbitt"
date: "2/24/2026"
output: github_document
---

### Load packages and data

```{r load-packages, message=FALSE}
library(tidyverse)
library(dplyr)
```

```{r}
kansas <- read_csv("/Users/georgenesbitt/Desktop/Data Science/Labs/lab_07_datascience/kansas_grouped_rolling_avg.csv")
```

### Exercise 1

```{r exploring misleading graph}
glimpse(kansas)
```

Below is the misleading graph:
### Excercise 2
```{r new less misleading graph}
ggplot(kansas, aes(x = date, 
                   y = rolling_avg, 
                   color = mask_mandate)) +
  geom_line() +
  scale_color_manual(values = c("Mask" = "orange",
                                "No Mask" = "blue")) +
  labs(
    title = "Kansas COVID-19 7-Day Rolling Average of Daily Cases Per 100K Population",
    subtitle = "Mask Counties vs. No-Mask Mandate Counties",
    caption = "Source: Kansas Department of Health and Environment",
    x = NULL,
    y = "Cases per 100K Population",
    color = NULL)
```
### Exercise 3
The message that is more clear in this graph than in the misleading one is the comparison of the dependent variable across mask conditions. Initially, it made the two look similar, however after equalizing the dependent variable along the y axis, you can see that mask counties had drastically higher numbers of cases. 

### Excercise 4
What this graph tells us (without any other context whatsoever) is that counties that have mask mandates also tend to have a higher number of covid cases. This does not reflect on what I know about mask wearing, however if you consider many factors, this might be due to that fact that counties with higher population density might also be more likely to have mask mandates in general. 

### Excercise 5
As I stated above, what likely contributes to this trend is that liberal cities typically have higher population densities and also typically have larger/more strict mask mandates. The lack of mask mandates in more rural areas with much lower population density is likely not enough to allow more covid cases.

### Excercise 6
For the this visual, I think that if I made the y axis percent change overtime as opposed to total case numbers, it might show a more balanced graph.

### Excercise 7
```{r create new variable index}
kansas <- kansas %>% group_by(mask_mandate) %>%
  mutate(indexed= (rolling_avg/first(rolling_avg))*100
)
```
Above I just created a new variable in kansas that is titled indexed (the percent change over time). Below I'll use the same code as the original graph with a new y variable (indexed)
```{r}
ggplot(kansas, aes(x = date, 
                   y = indexed, 
                   color = mask_mandate)) +
  geom_line() +
  scale_color_manual(values = c("Mask" = "orange",
                                "No Mask" = "blue")) +
  labs(
    title = "Kansas COVID-19 Percent change over time Per 100K Population",
    subtitle = "Mask Counties vs. No-Mask Mandate Counties",
    caption = "Source: Kansas Department of Health and Environment",
    x = NULL,
    y = "Percent of initial cases",
    color = NULL)
```

This shows that over time, the masked group fell far below the original number whereas the no mask group only fell about 2% below, essentially maintaining it's original numbers. 

### Excercise 8
